A Component-Based Sensor Network for \\ Environmental Monitoring

### 2009-03-13 docs/icccn2009/paper.tex
Edited conclusions
git-svn-id: svn://localhost/netbeams/trunk@376 559edde2-a4d4-44e0-96fc-4035e40d7396

The work described in this paper is a collaborative effort between the
Computer Science Department at San Francisco State University and the
Romberg Tiburon Center (RTC), a research institute focusing on the
understanding of complex marine and estuarine environments. The result
of this collaboration is NetBEAMS (Networked Bay Environmental
Assessment Monitoring System), an environmental sensor network
deployed in the San Francisco Bay. NetBEAMS provides means to program
and interrogate sensors at the remote field locations and to transmit
collected measurements to a database, where the data is rapidly
processed, archived, and made available to potential users in near
real-time. In the following we give an overview of the NetBEAMS
architecture and implementation.  Section \ref{SEC_BACKGROUND}
describes a typical environmental sensor used by the RTC. Section
\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
to efficiently build up an end-to-end environmental sensor network.
Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
of the DSP. In Section \ref{SEC_CONCLUSION} we provide a conclusion
and outlook.

@@

\section{Conclusions and Outlook}
\label{SEC_CONCLUSION}

The California coastal region is populated by a vast number of
remotely deployed sensors that monitor environmental conditions
ranging from weather to water quality to ocean surface currents.
These sensors are operated by data providers, including resource
management groups, research scientists, municipalities, and state and
federal agencies, who could use near real-time data streams to inform
decision that impact public health, public safety, and the protection
of resources.  In addition, there are efforts underway at the state
and federal level to collate the data collected by the individual data
providers into integrated monitoring networks that will enhance the
distribution of monitoring data to public, private, and federal users
groups.  Success of these efforts, whether small scale or a large
integrated network, depends on the first step in the data management
process: reliably obtaining data in near real-time from remotely
located sensors.  This step can be the weak link in the data
management pathway when telemetry options are unreliable.  The
availability of a plug-and-play computing platform designed to
communicate with a variety of sensor types and to provide data
telemetry to distant brick and mortar data centers would stabilize
data streams with poor telemetry and allowing expansion of sensor
coverage into regions not presently served.  The low power consumption
and low cost of these units would be particularly beneficial for
applications with long service cycles and for data providers trying to
maintain data streams on a limited budget.

>>>>

The work described in this paper is a collaborative effort between the
Computer Science Department at San Francisco State University and the
Romberg Tiburon Center (RTC), a research institute focusing on the
understanding of complex marine and estuarine environments. The result
of this collaboration is NetBEAMS (Networked Bay Environmental
Assessment Monitoring System), an environmental sensor network
deployed in the San Francisco Bay. NetBEAMS provides means to program
and interrogate sensors at the remote field locations and to transmit
collected measurements to a database, where the data is rapidly
processed, archived, and made available to potential users in near
real-time. In the following we give an overview of the NetBEAMS
architecture and implementation.  Section \ref{SEC_BACKGROUND}
describes a typical environmental sensor used by the RTC. Section
\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
to efficiently build up an end-to-end environmental sensor network.
Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
of the DSP. Section \ref{SEC_CONCLUSION} concludes this paper.

@@

\section{Conclusions}
\label{SEC_CONCLUSION}

The California coastal region is populated by a vast number of
remotely deployed sensors that monitor environmental conditions
ranging from weather to water quality to ocean surface currents.
These sensors are operated by data providers, including resource
management groups, research scientists, municipalities, and state and
federal agencies, who could use near real-time data streams to inform
decision that impact public health, public safety, and the protection
of resources. Success of these efforts, whether small scale or a large
integrated network, depends on the first step in the data management
process: reliably obtaining data in near real-time from remotely
located sensors.  This step can be the weak link in the data
management pathway when telemetry options are unreliable.  The
availability of a plug-and-play computing platform designed to
communicate with a variety of sensor types would stabilize data
streams with poor telemetry and allowing expansion of sensor coverage
into regions not presently served.

In this paper we introduce a practical application of a sensor network
in the domain of environmental monitoring. We have built an
environmental sensor network in the San Francisco Bay using highly
specialized, off-the-shelf water quality sensors. The resulting
infrastructure enables near-realtime access to sensor data collected
from remote locations. Our end-to-end solution furthermore allows
configuration and management of remotely deployed sensors. The key to
our infrastructure is the DSP; a micro-kernel plug-and-play
architecure that can be configured flexibly to specific needs through
special purpose components. The NetBEAMS infrastructure as described
in the paper is available as Open Source and can be viewed in
operation at \texttt{www.netbeams.org}.
